---
title: "Reproducible Research Project 1"
author: "Benjamin Milks"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Reading the Data

```{r}
unzip("rrproj1data.zip")
df <- read.csv("activity.csv")
```

## Manipulating the Data

```{r}
datelist <- levels(factor(df$date))
rdatelist <- lapply(datelist, strptime, format = "%Y-%m-%d")
rdlist <- vector()
for (i in 1:length(rdatelist)){
        ww <- rdatelist[[i]]
        rdlistlist <- append(rdlist, ww)
}
datesums <- vector()
for (i in datelist) { 
    qq <- sum(df$steps[df$date == i], na.rm = TRUE)
    datesums <- append(datesums, qq)}
```

## Histogram of Step Frequency 
 
```{r}
hist(datesums, breaks = 10, main = "Histogram of Step Frequency", xlab = "Steps")
```

## Median and Mean Steps in a Day

For my calculation here I remove the days where there were only NA's. The days that only had NA's were recorded as NA in my making of datesums.
```{r}
meanspd <- mean(datesums, na.rm = TRUE)
print(meanspd)
medianspd <- median(datesums, na.rm = TRUE)
print(medianspd)
```

## Looking at Number of Steps By Time of Day

```{r}
intervalmeans <- vector()
intervallist <- levels(factor(df$interval))
for (i in intervallist) { 
    pp <- mean(df$steps[df$interval == i], na.rm = TRUE)
    intervalmeans <- append(intervalmeans, pp)}

plot(intervallist, intervalmeans, type = "l", xlab = "Time Interval", ylab = "Mean Steps")
highstepinterval <- which.max(intervalmeans)
highstepinterval <- highstepinterval*5 ## Each spot on the list represents five minutes, so if we multiply the number we got by five, we know the number of minutes after midnight.
highstepinterval
## We may want to know what time this is.
hours <- highstepinterval/60
hours
## 8.666667 means 8 and 2/3 hours, or 8:40am.

## We know which interval has the highest mean, now we want to know what the maximum value is for mean steps during the 5 minutes.
max(intervalmeans)

```



# Strategy for Replacing Missing Data


## How Many Times is the Number of Steps Recorded as "na"?

```{r}
totalnas <- sum(is.na(df$steps))
print(totalnas)
```

## What to do to Replace the Missing Data

I think a good replacement would be to replace each na with the mean from the interval. I have a list of interval means from before. So to index the list I have to take the interval, say 25, and divide it by five then add one to get the correct spot from my list of interval means that was calculated earlier.
```{r}
dfnona <- df
m <- c(1:288)
mm <- 5*m
newiv <- rep(mm, 61)
dfnona<- cbind(dfnona, newiv)
head(dfnona)
for (i in 1:nrow(dfnona)){
        im <- dfnona[i,4]
        iint <- im/5
        imeanmarker <- iint 
        dfnona[i,1][is.na(dfnona[i,1])] <- intervalmeans[imeanmarker]
        
}
datelist <- levels(factor(dfnona$date))
rdatelist <- lapply(datelist, strptime, format = "%Y-%m-%d")
rdlist <- vector()
for (i in 1:length(rdatelist)){
        ww <- rdatelist[[i]]
        rdlistlist <- append(rdlist, ww)
}
datesums <- vector()
for (i in datelist) { 
    qq <- sum(dfnona$steps[dfnona$date == i], na.rm = TRUE)
    datesums <- append(datesums, qq)}
print(datesums)

hist(datesums, breaks = 10, main = "Histogram of Step Frequency After Impution", xlab = "Steps")

```


## Comparing weekends to weekdays to see if steps per day changes

```{r, echo=FALSE}
library(timeDate)
```


```{r}
weekenddf <- df
for (i in 1:nrow(weekenddf)){
        weekenddf[i,3]<- isWeekday(weekenddf[i,2])
        
        
}
weekenddf<- cbind(weekenddf, newiv)
library(lattice)
xyplot(steps~newiv | factor(interval), data = weekenddf, main = "Steps Per Interval Throughout the Day, Comparing Weekend(0) to Weekdays (1)", xlab = "Time Interval", ylab = "Number of Steps", alpha = 1/4 )
```
